### 2023-03-12 outputs/paper/paper.qmd
data cleaned and graphs created

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(ggplot2)
library(haven)
library(janitor)
library(tidyr)
library(dplyr)
```

```{r}
#| label: fig-1
#| fig-cap: Highest Degree of Education Obtained by the Respondant.
#| echo: false
#| warning: false
#| message: false

clean_data1 <- read_csv("~/Documents/GSS_education/inputs/data/clean_data1.csv")

clean_data1 |> 
  mutate(education = fct_relevel(education, 
            "< High School", "High School", "Some College", "College", "Advanced Degree")) |>
  ggplot(aes(x = education)) +
  geom_bar(fill = "slateblue1") +
  theme_minimal() +
  ylim(0, 2000) +
  labs(x = "Achieved Education Level",
       y = "Number of Participants",
       title = "Education Level of the Respondant.")
```

```{r}
#| label: fig-2
#| fig-cap: Highest Degree of Education Obtained by the Respondants' Mother.
#| echo: false
#| warning: false
#| message: false

clean_data2 <- read_csv("~/Documents/GSS_education/inputs/data/clean_data2.csv")

clean_data2 |> 
  mutate(education = fct_relevel(education, 
            "< High School", "High School", "Some College", "College", "Advanced Degree")) |>
  ggplot(aes(x = education)) +
  geom_bar(fill = "violet") +
  theme_minimal() +
  ylim(0, 2000) +
  labs(x = "Achieved Education Level",
       y = "Number of Participants",
       title = "Education Level of Respondants' Mother.")
```

```{r}
#| label: fig-3
#| fig-cap: Highest Degree of Education Obtained by the Respondants' Father.
#| echo: false
#| warning: false
#| message: false

clean_data3 <- read_csv("~/Documents/GSS_education/inputs/data/clean_data3.csv")

clean_data3 |> 
  mutate(education = fct_relevel(education, 
            "< High School", "High School", "Some College", "College", "Advanced Degree")) |>
  ggplot(aes(x = education)) +
  geom_bar(fill = "royalblue1") +
  theme_minimal() +
  ylim(0, 2000) +
  labs(x = "Achieved Education Level",
       y = "Number of Participants",
       title = "Education Level of Respondants' Father.")
```

# Introduction

# Data {#sec-data}

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References